---
title: "dplyr"
author: "Syed Faraz Ali"
date: "9/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing libraries 

```{r libraries, message = FALSE}
library(dplyr)
library(ggplot2)
library(data.table)
```

### Setting Working Directory

```{r}
setwd("~/Downloads")
getwd()
```

### Reading the data 

```{r cache = TRUE}
flights <- tbl_df(read.csv("flights.csv", stringsAsFactors = FALSE))
flights$date <- as.Date(flights$date)

weather <- tbl_df(read.csv("weather.csv", stringsAsFactors = FALSE))
weather$date <- as.Date(weather$date)

planes <- tbl_df(read.csv("planes.csv", stringsAsFactors = FALSE))

airports <- tbl_df(read.csv("airports.csv", stringsAsFactors = FALSE))
```


### Printing a portion of all the data 

```{r}
head(flights)
head(weather)
head(planes)
head(airports)
```


```{r}
summary(flights)
```


### The Five Verbs 

* Filter : Keep rows matching criteria 
* Select : Pick columns by name 
* Arrange : Reorder rows 
* Mutate : Add new variables 
* Summarise : Reduce Variables 


```{r}
df <- data.frame(color = c("blue", "black","blue", "blue", "black"),
                 value = 1:5)
df
```

Using filter 

```{r}
filter (df, color == "blue")

filter (df , value %in% c(3,4))
```


### Practicing filter 

All the flights to SFO or OAK 

```{r}
sfo <- filter(flights, dest %in% c("SFO", "OAK"))
print(sfo)
```

All the flights in January 

```{r}
jan <- filter(flights, month(flights$date) == 1)
print(jan)
```

Flights that departed between midnight and 5 AM 

```{r}
filter(flights, hour >=0 & hour <= 5)
```

Flights that got delayed by more than hour 

```{r}
filter(flights, arr_delay > 60 )
```


Print out df 

```{r}
print(df)
```

Using select : select() keeps only the variables you mention; 

```{r}
select(df , value)
```

















